Review of Edge Detectors

### 2014-11-11 old_versions/edge-detection-review-v12.tex
second round of third revision, improving references

%-------------------------------------------------------------------------------
% A Review of Classic Edge Detectors - v10
% by Haldo Spontón & Juan Cardelino
% IPOL 2012
%-------------------------------------------------------------------------------

\documentclass{ipol}
\ipolSetYear{2012}
\ipolSetMonth{06}
\ipolSetDay{06}
%\ipolSetID{arced}
\ipolSetTitle{A Review of Classic Edge Detectors}
\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$, Rafael Grompone von Gioi$^3$}
\ipolSetAffiliations{%
$^1$ IIE, UdelaR, Uruguay (\texttt{haldos@fing.edu.uy})\\
$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})\\
$^3$ CMLA, ENS Cachan, France (\texttt{grompone@cmla.ens-cachan.fr})}

@@

% cambiar lo de segmenting images. Decir que edge detection es una de las operaciones
% basicas y mas antiguas en el procesamiento de imagenes.
% Edge detection is one of the first approaches published in the literature for 
% segmenting images. 
The basic idea of edge detection algorithms is to detect abrupt changes in image intensity. 
Detecting those changes in intensity for the purpose of finding edges in images 
can be accomplished using first or second order derivatives.

@@

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. %But when finding 
%diagonal edges is desired, 2-D masks are needed. 
The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal 
diferences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
usual to scale the mask values, in order to have unit $L_2$ norm.

@@

The simplest digital approximation to the partial derivatives using masks of size $3\times3$ are obtained 
by taking the difference between the third and first rows (or columns) of the $3\times3$ region. The
difference between the third and first rows approximates the derivative in the x-direction, and 
the difference between the third and first columns approximates the derivative in the y-direction.
These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
These masks are called \textit{Prewitt operators}.

\begin{SCfigure}[1][!h]
	\centering
	\begin{squarecells}{3}
		$-\frac{1}{6}$ 	& $-\frac{1}{6}$ 	& $-\frac{1}{6}$	\nline
		0 			& 0			& 0			\nline
		$\frac{1}{6}$ 	& $\frac{1}{6}$ 	& $\frac{1}{6}$	\nline
	\end{squarecells}
	\quad
	\begin{squarecells}{3}
		$-\frac{1}{6}$ 	& 0 	& $\frac{1}{6}$	\nline
		$-\frac{1}{6}$	& 0	& $\frac{1}{6}$	\nline
		$-\frac{1}{6}$ 	& 0 	& $\frac{1}{6}$	\nline
	\end{squarecells}
	\caption{Normalized \textit{Prewitt} 2-D masks of size $3\times3$.}
	\label{fig:prewitt}
\end{SCfigure}

@@

A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
difference. It can be shown that using double weight in the center location provides better image smoothing. This
variation is implemented using masks in Figure \ref{fig:sobel}. These operators are called 
\textit{Sobel operators}.

@@

\begin{SCfigure}[1][!h]
	\centering
	\begin{squarecells}{3}
		$-\frac{1}{8}$ 	& $-\frac{1}{4}$ 	& $-\frac{1}{8}$	\nline
		0 			& 0			& 0			\nline
		$\frac{1}{8}$ 	& $\frac{1}{4}$ 	& $\frac{1}{8}$	\nline
	\end{squarecells}
	\quad
	\begin{squarecells}{3}approximate
		$-\frac{1}{8}$ 	& 0 	& $\frac{1}{8}$	\nline
		$-\frac{1}{4}$	& 0	& $\frac{1}{4}$	\nline
		$-\frac{1}{8}$ 	& 0 	& $\frac{1}{8}$	\nline
	\end{squarecells}
	\caption{Normalized \textit{Sobel} 2-D masks of size $3\times3$.}
	\label{fig:sobel}
\end{SCfigure}

The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
use Sobel operators, because they are better localized, and their filter are also less aliased, 
because of the weighted shape of $[1\ 2\ 1]$.

@@

Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
\begin{equation*}
	\mbox{edges}[i,j] = \begin{cases} 0,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
									1,& \mbox{if}\ M[i,j]<\mbox{th}
					  \end{cases}
\end{equation*}approximate
where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}).

@@

Then, the least squares problem can be approximately solved by computing the pseudo inverse of $\mathbf{A}$,
\begin{equation*}
	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f} .\\
\end{equation*}
$\mathbf{B}$ is a $10\times25$ matrix, 
\begin{equation*}
	\begin{bmatrix}
		b_{1,1}		& b_{1,2}	& \cdots	& b_{1,25}	\\
		b_{2,1}		& b_{2,2}	& \cdots	& b_{2,25}	\\
		\vdots		& \vdots	& \ddots	& \vdots	\\
		b_{10,1}	& b_{10,2}	& \cdots	& b_{10,25}
	\end{bmatrix}
	\times
	\begin{bmatrix}
		f_1		\\
		f_2		\\
		\vdots	\\
		f_{25}
	\end{bmatrix}
	=
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix} .\\
\end{equation*}
For each coefficient ($i$ from $1$ to $10$), 
\begin{equation}
	\label{eq:coefficients}
	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \cdots + b_{i,25}f_{25}% \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
\end{equation}
where
\begin{equation}
	\label{eq:b_i}
	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \cdots	& b_{i,5}	\\
									b_{i,6}		& b_{i,7}	& \cdots	& b_{i,10}	\\
									\vdots		& \vdots	& \ddots	& \vdots	\\
									b_{i,21}	& b_{i,22}	& \cdots	& b_{i,25}	\\
					\end{bmatrix}.\\
\end{equation}
%% VER UNA FORMA DE ESCRIBIR TAMBIEN LOS f1...f25 PARA MOSTRAR POR QUE LOS bi,1 ... bi,25 ESTAN DISPUESTOS COMO UNA MATRIZ. ES PORQUE LOS f SON UN ENTORNO 5X5 DEL PIXEL, Y SE MULTIPLIAN PUNTO A PUNTO POR LOS b.

@@

The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If just the first order terms of this 
polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approximated as
\begin{align}
	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}.  
\label{eq:sincos}
\end{align}
Now substituting the variables $x$ and $y$ in polar form as
\begin{equation*}
	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} 
\end{equation*}
in the bi-cubic polynomial, we obtain
\begin{equation}
	f_{\theta}(\rho) = C_0 + C_1\rho + C_2\rho^2 + C_3\rho^3 ,
\end{equation}
where
\begin{align}
\label{eq:c}
	C_0 & = k_1 \nonumber \nonumber \\
	C_1 & = k_2\sin(\theta) + k_3\cos(\theta) \nonumber \\
	C_2 & = k_4\sin^2(\theta) + k_5\sin(\theta)\cos(\theta) + k_6\cos^2(\theta) \nonumber \\
	C_3 & = k_7\sin^3(\theta) + k_8\sin^2(\theta)\cos(\theta) + k_9\sin(\theta)\cos^2(\theta) + k_{10}\cos^3(\theta). 
\end{align}
The derivatives are obtained as follows:
\begin{align}
	f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \nonumber \\
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
	f'''_{\theta}(\rho) = 6C_3 .\nonumber 
\end{align}
Then, the condition that the second derivative is equal to zero becomes
\begin{equation}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0,
\end{equation}
where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
should be zero at $\rho=0$, but this condition is relaxed, allowing the polynomial to become zero in a neighborhood 
of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
value is between $0.4$ and $0.6$.

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\STATE $im \leftarrow$ input image
\STATE Define the ten $5x5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ $\sum_{n,m}neighbors[n,m]mask_j[n,m]$ % convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of n is 
large enough to have a good approximation of the Gaussian function in the kernel.}
	\label{fig:gaussian_kernel}
\end{SCfigure}

@@

The Laplacian of Gaussian $\nabla^2G(x,y)$ can be obtained analyically first and then a discrete mask 
can be computed by sampling the analytical kernel. Using this kernel for edge detection involves only 
one convolution with the input image (unlike the case of Gaussian kernel, in which two convolutions 
have to be performed, one with the kernel and another with the Laplacian operator), but the kernel 
support must be greater in order to obtain the same precision..

@@

\begin{equation} 
	\frac{\partial}{\partial x}G(x,y)=-\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
Therefore we obtain
\begin{equation}
	\label{eq:log_function}
	LoG(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
\end{equation}\\

@@

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of n is sufficient 
to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
Gaussian kernel.}
	\label{fig:log_kernel}
\end{SCfigure}

@@

When making a convolution, it is necessary to define the boundary conditions used to calculate such convolution 
around the edges of the image, and to get a valid output the same size as input image. In this paper, 
two methods were implemented: zero-padding and boundary reflection. Zero-padding implies complete 
the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
relative to the edge of the image. Direct convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution or recursive filtering\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.

@@

The results of each algorithm are first shown in a simple case, where a test image composed of a white square on a black background (shown in Figure \ref{fig:original1}) is used.

@@

Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm. The execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop.} for each algorithm are shown in Table \ref{exectime1}.

\begin{figure}[h!]
	\centering
	\subfigure[Roberts. $th=0.1$.]{\label{fig:result1-a}\includegraphics[width=0.2\textwidth]{results/square127_roberts.png}}
	\quad
	\subfigure[Prewitt. $th=0.1$.]{\label{fig:result1-b}\includegraphics[width=0.2\textwidth]{results/square127_prewitt.png}}
	\quad
	\subfigure[Sobel. $th=0.1$.]{\label{fig:result1-c}\includegraphics[width=0.2\textwidth]{results/square127_sobel.png}}

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image. The first three (\subref{fig:result1-a}, \subref{fig:result1-b} and \subref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\subref{fig:result1-d} and \subref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\subref{fig:result1-f}) corresponds to the Haralick algorithm. Note the rounded corners on the Marr-Hildreth results, due to Gaussian blur.}
	\label{fig:result1}
\end{figure}

\begin{table}[h!]
	\begin{center}
	\begin{tabular}{| l | r |}
		\hline \rule{0pt}{3ex}
		\cellcolor[gray]{0.8} \textbf{Algorithm}	& \cellcolor[gray]{0.8} \textbf{Execution time (s)}	\\ \hline \rule{0pt}{3ex}
		Roberts, Prewitt and Sobel					& $0.020 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (Gaussian)					& $0.030 \ s$										\\ \hline \rule{0pt}{3ex}
		Marr-Hildreth (LoG)							& $0.050 \ s$										\\ \hline \rule{0pt}{3ex}
		Haralick									& $0.040 \ s$										\\
		\hline
	\end{tabular}
	\end{center}
	\caption{Execution time of the algorithms (including I/O) using the Figure \ref{fig:original1} as input image (127$\times$127 pixels, 1 channel).}
	\label{exectime1}
\end{table}
\vspace{0.5cm}

It is appreciated that in the simplest cases, the first derivative edge detection algorithms are running better and faster. All algorithms show consistent results. Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm.

Now the performance of each algorithm is analyzed, using a natural image shown in Figure \ref{fig:original2}.

@@

The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since more convolutions (and with larger kernels) are performed. In real images, they have better behavior than first derivative algorithms.

@@

\bibliography{bibliography}

>>>>

%-------------------------------------------------------------------------------
% A Review of Classic Edge Detectors - v12
% by Haldo Spontón & Juan Cardelino
% IPOL 2012
%-------------------------------------------------------------------------------

\documentclass{ipol}
\ipolSetYear{2012}
\ipolSetMonth{06}
\ipolSetDay{06}
%\ipolSetID{arced}
\ipolSetTitle{A Review of Classic Edge Detectors}
\ipolSetAuthors{Haldo Spont\'on$^1$, Juan Cardelino$^2$}
\ipolSetAffiliations{%
$^1$ IIE, UdelaR, Uruguay (\texttt{haldos@fing.edu.uy})\\
$^2$ IIE, UdelaR, Uruguay (\texttt{juanc@fing.edu.uy})}

@@

% cambiar lo de segmenting images. Decir que edge detection es una de las operaciones
% basicas y mas antiguas en el procesamiento de imagenes.
% Edge detection is one of the first approaches published in the literature for 
% segmenting images. 
The basic idea of edge detection algorithms is to detect abrupt changes in image intensity. 
Detecting those changes can be accomplished using first or second order derivatives.

@@

Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$
by filtering the image $f(x,y)$ with the 1-D masks in Figure \ref{fig:1dmasks}. %But when finding 
%diagonal edges is desired, 2-D masks are needed. 
The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal 
diferences implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
usual to scale the mask values, in order to have unit derivative when unitary steps in intensity occurs.

@@

The simplest digital approximation to the partial derivatives using masks of size $3\times3$ are obtained 
by taking the difference between the third and first rows (or columns) of the $3\times3$ region. The
difference between the third and first rows approximates the derivative in the x-direction, and 
the difference between the third and first columns approximates the derivative in the y-direction.
These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
These masks are called \textit{Prewitt operators} \cite{prewitt1970object}.

\begin{SCfigure}[1][!h]
	\centering
	\begin{squarecells}{3}
		$-\frac{1}{3}$ 	& $-\frac{1}{3}$ 	& $-\frac{1}{3}$	\nline
		0 			& 0			& 0			\nline
		$\frac{1}{3}$ 	& $\frac{1}{3}$ 	& $\frac{1}{3}$	\nline
	\end{squarecells}
	\quad
	\begin{squarecells}{3}
		$-\frac{1}{3}$ 	& 0 	& $\frac{1}{3}$	\nline
		$-\frac{1}{3}$	& 0	& $\frac{1}{3}$	\nline
		$-\frac{1}{3}$ 	& 0 	& $\frac{1}{3}$	\nline
	\end{squarecells}
	\caption{Normalized \textit{Prewitt} 2-D masks of size $3\times3$.}
	\label{fig:prewitt}
\end{SCfigure}

@@

A slight variation of the Prewitt operators uses more weight on the central coefficients of the 
difference. It can be shown that using double weight in the center location provides better image smoothing. This
variation is implemented using the masks in Figure \ref{fig:sobel}. These operators are called 
\textit{Sobel operators}.

@@

\begin{SCfigure}[1][!h]
	\centering
	\begin{squarecells}{3}
		$-\frac{1}{4}$ 	& $-\frac{1}{2}$ 	& $-\frac{1}{4}$	\nline
		0 			& 0			& 0			\nline
		$\frac{1}{4}$ 	& $\frac{1}{2}$ 	& $\frac{1}{4}$	\nline
	\end{squarecells}
	\quad
	\begin{squarecells}{3}
		$-\frac{1}{4}$ 	& 0 	& $\frac{1}{4}$	\nline
		$-\frac{1}{2}$	& 0	& $\frac{1}{2}$	\nline
		$-\frac{1}{4}$ 	& 0 	& $\frac{1}{4}$	\nline
	\end{squarecells}
	\caption{\textit{Sobel} 2-D masks of size $3\times3$.}
	\label{fig:sobel}
\end{SCfigure}

The computational difference between Prewitt and Sobel masks is not important, thus It is preferable to 
use Sobel operators, because edges are better localized, and their filter are also less aliased, 
because of the weighted shape of $[1\ 2\ 1]$.

@@

Finally, the edges image is obtained by thresholding the gradient magnitude image, i.e.\ for pixel at position $i$ in the image compute
\begin{equation*}
	\mbox{edges}[i,j] = \begin{cases} 1,& \mbox{if}\ M[i,j]\geq\mbox{th} \\
									0,& \mbox{if}\ M[i,j]<\mbox{th}
					  \end{cases}
\end{equation*}approximate
where $\mbox{th}$ is the threshold. A black and white image is obtained, in which edge points are indicated in white (see Figure \ref{fig:thresholding}).

@@

Then, the least squares problem can be approximately solved by computing the pseudo inverse of $\mathbf{A}$,
\begin{equation*}
	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f} .\\
\end{equation*}
$\mathbf{B}$ is a $10\times25$ matrix, 
\begin{equation*}
	\begin{bmatrix}
		b_{1,1}		& b_{1,2}	& \cdots	& b_{1,25}	\\
		b_{2,1}		& b_{2,2}	& \cdots	& b_{2,25}	\\
		\vdots		& \vdots	& \ddots	& \vdots	\\
		b_{10,1}	& b_{10,2}	& \cdots	& b_{10,25}
	\end{bmatrix}
	\times
	\begin{bmatrix}
		f_1		\\
		f_2		\\
		\vdots	\\
		f_{25}
	\end{bmatrix}
	=
	\begin{bmatrix}
		k_1		\\
		k_2		\\
		\vdots	\\
		k_{10}
	\end{bmatrix} .\\
\end{equation*}
For each coefficient ($i$ from $1$ to $10$), 
\begin{equation}
	\label{eq:coefficients}
	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \cdots + b_{i,25}f_{25}% \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
\end{equation}
where\footnote{In this matrix the coefficients $b_i$ are simply spatially arranged with the same topology as the pixels they multiply.}
\begin{equation}
	\label{eq:b_i}
	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \cdots	& b_{i,5}	\\
									b_{i,6}		& b_{i,7}	& \cdots	& b_{i,10}	\\
									\vdots		& \vdots	& \ddots	& \vdots	\\
									b_{i,21}	& b_{i,22}	& \cdots	& b_{i,25}	\\
					\end{bmatrix}.\\
\end{equation}
%% VER UNA FORMA DE ESCRIBIR TAMBIEN LOS f1...f25 PARA MOSTRAR POR QUE LOS bi,1 ... bi,25 ESTAN DISPUESTOS COMO UNA MATRIZ. ES PORQUE LOS f SON UN ENTORNO 5X5 DEL PIXEL, Y SE MULTIPLIAN PUNTO A PUNTO POR LOS b.

@@

The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If just the first order terms of this 
polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approximated as
\begin{align}
	\sin(\theta) & = -\frac{k_2}{\sqrt{k_2^2 + k_3^2}} \nonumber \\
	\cos(\theta) & = -\frac{k_3}{\sqrt{k_2^2 + k_3^2}}.  
\label{eq:sincos}
\end{align}
Now substituting the variables $x$ and $y$ in polar form as
\begin{equation*}
	x = \rho\cos{\theta} \: \text{and} \: y = \rho\sin{\theta} 
\end{equation*}
in the bi-cubic polynomial, we obtain
\begin{equation}
	f_{\theta}(\rho) = C_0 + C_1\rho + C_2\rho^2 + C_3\rho^3 ,
\end{equation}
where
\begin{align}
\label{eq:c}
	C_0 & = k_1 \nonumber \nonumber \\
	C_1 & = k_2\sin(\theta) + k_3\cos(\theta) \nonumber \\
	C_2 & = k_4\sin^2(\theta) + k_5\sin(\theta)\cos(\theta) + k_6\cos^2(\theta) \nonumber \\
	C_3 & = k_7\sin^3(\theta) + k_8\sin^2(\theta)\cos(\theta) + k_9\sin(\theta)\cos^2(\theta) + k_{10}\cos^3(\theta). 
\end{align}
The derivatives are obtained as follows:
\begin{align}
	f'_{\theta}(\rho) = C_1 + 2C_2\rho + 3C_3\rho^2 \nonumber \\
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho \nonumber \\
	f'''_{\theta}(\rho) = 6C_3 .\nonumber 
\end{align}
Then, the condition that the second derivative is equal to zero becomes
\begin{equation}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0,
\end{equation}
where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
should be zero at $\rho=0$, but this condition is relaxed\footnote{This polynomial being zero at $\rho = 0$ means that the edge passes right through the exact coordinates of the pixel. Since the edge may be slightly displaced, other $\rho$ values are permitted as long as the edge is inside the pixel, and for this, the maximum distance from the center of the pixel may be $1/\sqrt{2}$.}, allowing the polynomial to become zero in a neighborhood 
of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\frac{1}{\sqrt{2}}$. Its optimal 
value is between $0.4$ and $0.6$.

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.BibTex | RIS | RefWorks
\STATE $im \leftarrow$ input image
\STATE Define the ten $5x5$ masks ($mask_1\dots mask_{10}$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ $\sum_{n,m}neighbors[n,m]mask_j[n,m]$ % convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

\begin{SCfigure}[][!t]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_gaussian.pdf}
	\caption{Gaussian kernel, $\sigma=4$, $n=25$. Is easy to see that the selected value of $n$ is 
large enough to have a good approximation of the Gaussian function in the kernel.}
	\label{fig:gaussian_kernel}
\end{SCfigure}

@@

The Laplacian of Gaussian $\nabla^2G(x,y)$ can be obtained analyically first and then a discrete mask 
can be computed by sampling the analytical kernel. Using this kernel for edge detection involves only 
one convolution with the input image (unlike the case of Gaussian kernel, in which two convolutions 
have to be performed, one with the kernel and another with the Laplacian operator), but the kernel 
support must be greater in order to obtain the same precision.

@@

\begin{equation} 
	\frac{\partial}{\partial x}G(x,y)=-\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial}{\partial y}G(x,y)=-\frac{1}{\sqrt{2\pi\sigma^2}}\frac{y}{\sigma^2}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 x}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2},
\end{equation}
\begin{equation} 
	\frac{\partial^2}{\partial^2 y}G(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{y^2-\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
\end{equation}
Therefore we obtain
\begin{equation}
	\label{eq:log_function}
	LoG(x,y)=\frac{1}{\sqrt{2\pi\sigma^2}}\frac{x^2+y^2-2\sigma^2}{\sigma^4}e^{-(x^2+y^2)/2\sigma^2}.
\end{equation}\\

@@

\begin{SCfigure}[][!b]
	\centering
	\includegraphics[width=0.5\textwidth]{kernel_log.pdf}
	\caption{Laplacian of a Gaussian kernel, $\sigma=4$, $n=31$. The selected value of $n$ is sufficient 
to have a good approximation of the LoG function in the kernel, but is greater than in the case of 
Gaussian kernel.}
	\label{fig:log_kernel}
\end{SCfigure}

@@

When making a convolution, it is necessary to define the boundary conditions used to calculate such convolution 
around the edges of the image, and to get a valid output the same size as input image. In this paper, 
two methods were implemented: zero-padding and boundary reflection. Zero-padding implies complete 
the borders of the image with many zero valued pixels as needed (depending on the size of the convolution 
kernel). Reflection involves completing those pixels with the corresponding symmetric pixel value 
relative to the edge of the image. Direct convolution is not the only way of filtering an image with a kernel. There are other methods such as FFT convolution or recursive filtering.%\footnote{Implementations of these methods can be found in the \href{https://tools.ipol.im/wiki/author/code/hatchery/}{IPOL Code Hatchery}.}.

@@

The results of each algorithm are first shown for a simple image composed of a white square on a black background, see Figure \ref{fig:original1}.

@@

Figures \ref{fig:result1-a} to \ref{fig:result1-f} show the output of each algorithm (including execution times\footnote{Running on Intel Core i3 CPU (2.53GHz), 3 Gb RAM, standard laptop. Note that the execution time of the algorithms Roberts, Prewitt and Sobel are the same, because they run simultaneously.}).

\begin{figure}[h!]
	\centering
	\subfigure[Roberts. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-a}\includegraphics[width=0.2\textwidth]{results/square127_roberts.png}}
	\quad
	\subfigure[Prewitt. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-b}\includegraphics[width=0.2\textwidth]{results/square127_prewitt.png}}
	\quad
	\subfigure[Sobel. $th=0.1$. $\text{Exec-time}:0.02s$.]{\label{fig:result1-c}\includegraphics[width=0.2\textwidth]{results/square127_sobel.png}}

	\subfigure[Marr-Hildreth (Gaussian). $\sigma=1.5$, $n=13$, $th_{ZC}=0.1$. $\text{Exec-time}:0.03s$.]{\label{fig:result1-d}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth.png}}
	\quad
	\subfigure[Marr-Hildreth (LoG). $\sigma=1.5$, $n=17$, $th_{ZC}=0.1$. $\text{Exec-time}:0.05s$.]{\label{fig:result1-e}\includegraphics[width=0.2\textwidth]{results/square127_marr-hildreth-log.png}}
	\quad
	\subfigure[Haralick. $\rho=0.4$. $\text{Exec-time}:0.04s$.]{\label{fig:result1-f}\includegraphics[width=0.2\textwidth]{results/square127_haralick.png}}
	\caption{Results of the algorithms using the Figure \ref{fig:original1} as input image. The first three (\subref{fig:result1-a}, \subref{fig:result1-b} and \subref{fig:result1-c}) correspond to the first derivative methods (Roberts, Prewitt and Sobel), then the following two (\subref{fig:result1-d} and \subref{fig:result1-e}) are from the Marr-Hildreth algorithm using Gaussian and LoG kernels, and the last one (\subref{fig:result1-f}) corresponds to the Haralick algorithm. Note the rounded corners on the Marr-Hildreth results, due to Gaussian blur.}
	\label{fig:result1}
\end{figure}

In this simple case, the first derivative edge detection algorithms run better and faster. All algorithms show consistent results. Thicker edges appear in the results of Haralick, due to smooth image model that is imposed in this algorithm.

@@

The second derivative algoritms (Marr-Hildreth \& Haralick) involve several more operations, since convolutions with larger kernels are performed. In real images, they have better behavior than first derivative algorithms.

@@

\bibliographystyle{abbvr}
\bibliography{bibliography}
